---
title: "Options dans le cadre Black-Scholes"
subtitle: 'TP-2: Pricing Vanna-Volga'
date: "Version: `r format(Sys.Date(), '%d %b %Y')`"
output:
  pdf_document:
    keep_tex: yes
    fig_caption: yes
    latex_engine: pdflatex
  word_document: default
geometry: margin=1in
header-includes:
- \usepackage[utf8]{inputenc}
- \usepackage{amsmath}
- \usepackage{amsfonts}
- \usepackage{amssymb}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```
The purpose of this problem set is to explore the Vanna-Volga pricing model. In this problem set, you will use the following functions:

GBSPrice: Price of a vanilla option:
\[
P = f(\text{PutCall}, S, K, T, r, b, \sigma)
\]

where:
\begin{description}
\item[PutCall] 'c' for a call, 'p' for a put
\item[$b$] cost of carry: ridk free rate $r$ less dividend yield $d$
\item[$r$] risk-free rate
\end{description}

```{r, echo=TRUE}
GBSPrice <- function(PutCall, S, K, T, r, b, sigma) {
  d1 <- (log(S/K) + (b+sigma^2/2)*T)/(sigma*sqrt(T))
  d2 <- d1 - sigma*sqrt(T)

  if(PutCall == 'c')  
    px <- S*exp((b-r)*T)*pnorm(d1) - K*exp(-r*T)*pnorm(d2)
  else
    px <- K*exp(-r*T)*pnorm(-d2) - S*exp((b-r)*T)*pnorm(-d1)

px
}
```

GBSVega: Vega ($\frac{\partial P}{\partial \sigma}$) of a Vanilla option:

```{r, echo=TRUE}
GBSVega <- function(PutCall, S, K, T, r, b, sigma) {
  d1 <- (log(S/K) + (b+sigma^2/2)*T)/(sigma*sqrt(T))
  S*exp((b-r)*T) * dnorm(d1)
}
```

# Volatility Interpolation

Given the implied volatility at three strikes, we will use the Vanna-Volga pricing method to interpolate the
volatility curve. Assume $r=0, b=0, T=1, \mbox{Spot} = 100$.

```{r, echo=TRUE}
# Benchmark data: (strike, volatility)
VolData <- list(c(80, .32), c(100, .30), c(120, .315))
```

Let's first define an array of pricing functions for the 
benchmark instruments:

```{r, echo=TRUE}
C1 <- function(vol=sigma, spot=Spot) GBSPrice(PutCall='c', S=spot, K=VolData[[1]][1], T=T, r=r, b=b, sigma=vol)

C2 <- function(vol=sigma, spot=Spot) GBSPrice(PutCall='c', S=spot, K=VolData[[2]][1], T=T, r=r, b=b, sigma=vol)

C3 <- function(vol=sigma, spot=Spot) GBSPrice(PutCall='c', S=spot, K=VolData[[3]][1], T=T, r=r, b=b, sigma=vol)

C <- c(C1, C2, C3)
```

1. Write a utility functions to compute the risk indicators,
 all by finite difference:
 
```{r, echo=TRUE}
Vega <- function(f, vol, spot=Spot) {
  d_vol <- 10e-5
  return( (f(vol+d_vol*vol, Spot)-f(vol-d_vol*vol, Spot))/(2*d_vol) )
}

Vanna <- function(f, vol, spot=Spot) {
  d_vol <- 10e-5
  d_spot <- 10e-5
  return( (f(vol+d_vol*vol, Spot+d_spot*Spot)+
           f(vol-d_vol*vol, Spot-d_spot*Spot)-
           f(vol+d_vol*vol, Spot-d_spot*Spot)-
           f(vol-d_vol*vol, Spot+d_spot*Spot))/(4*d_vol*d_spot) )
}
  
Volga <- function(f, vol, spot=Spot) {
  d_vol <- 10e-5
  return( (f(vol+d_vol*vol, Spot) - 2*f(vol, Spot) + f(vol-d_vol*vol, Spot))/d_vol**2 )
  }
```

Then, the calculation of vega for the three benchmark options may be performed by:

```{r, echo=TRUE}
r<-0
b<-0
T<-1
Spot <- 100
B.vega <- sapply(1:3, function(i) Vega(C[[i]], VolData[[i]][2]))
```

2. Compute vectors of vega, vanna, volga for the three hedge instruments

```{r, echo=TRUE}
r<-0
b<-0
T<-1
Spot <- 100
B.vega.benchmark <- sapply(1:3, function(i) Vega(C[[i]], VolData[[i]][2]))
B.vanna.benchmark <- sapply(1:3, function(i) Vanna(C[[i]], VolData[[i]][2]))
B.volga.benchmark <- sapply(1:3, function(i) Volga(C[[i]], VolData[[i]][2]))
print("B.vega.benchmark for C1, C2 and C3")
B.vega.benchmark
print("B.vanna.benchmark for C1, C2 and C3")
B.vanna.benchmark
print("B.volga.benchmark for C1, C2 and C3")
B.volga.benchmark
```

3. Choose a new strike for which we want to compute the implied volatility.
Let's choose $K=110$.

4. Compute the risk indicators for a call option struck at that strike. 

```{r, echo=TRUE}
r<-0
b<-0
T<-1
Spot <- 100
K <- 110


VolData.ATM <- VolData[[2]][2]

f <- function(vol=VolData.ATM, spot=Spot){
  GBSPrice(PutCall='c', S=spot, K=K, T=T, r=r, b=b, sigma=vol)
}
B.vega <- Vega(f, VolData.ATM, Spot)
B.vanna <- Vanna(f, VolData.ATM, Spot)
B.volga <- Volga(f, VolData.ATM, Spot)
print("vega for K=110")
B.vega
print("vanna for K=110")
B.vanna
print("volga for K=110")
B.volga
b.risk <- c(B.vega, B.vanna, B.volga)
```
    
5. Compute the Vanna-Volga price adjustment and the corresponding implied volatility.
    
```{r, echo=TRUE}
A <- matrix(data = c(B.vega.benchmark, B.vanna.benchmark, B.volga.benchmark), nrow =3)
A <- t(A)
X <- solve(A, b.risk)
print("Matrice A =")
print(A)
print("Risk indicators (b) = ")
print(b.risk)
print("Weights =")
print(X)

vol.K <- VolData[[1]][2]*X[1] + VolData[[2]][2]*X[2] + VolData[[3]][2]*X[3]
print("vol de K = 110")
print(vol.K)

C1.M <- GBSPrice(PutCall='c', S=100, K=80, T=T, r=r, b=b, sigma=VolData.ATM)

C2.M <- GBSPrice(PutCall='c', S=100, K=100, T=T, r=r, b=b, sigma=VolData.ATM)

C3.M <- GBSPrice(PutCall='c', S=100, K=120, T=T, r=r, b=b, sigma=VolData.ATM)

C.M <- c(C1.M, C2.M, C3.M)

C1.BS <- GBSPrice(PutCall='c', S=100, K=80, T=T, r=r, b=b, sigma=VolData.ATM)

C2.BS <- GBSPrice(PutCall='c', S=100, K=100, T=T, r=r, b=b, sigma=VolData.ATM)

C3.BS <- GBSPrice(PutCall='c', S=100, K=120, T=T, r=r, b=b, sigma=VolData.ATM)

C.BS <- c(C1.BS, C2.BS, C3.BS)

O.BS <- GBSPrice(PutCall='c', S=100, K=110, T=T, r=r, b=b, sigma=VolData[[2]][2])


somme <- 0
for (i in 1:3) {
  somme <- somme + X[i]*(C.M[i]-C.BS[i])
}
O.M <- O.BS + somme
print("Price with BS")
print(O.BS)
print("Price ajusted")
print(O.M)

```




6. Wrap the above logic in a function in order to
interpolate/extrapolate the vol curve from $K=70$ to $K=130$

```{r, echo=F}

r<-0
b<-0
T<-1
Spot <- 100
K <- seq (70, 130 ,0.5)
vol.imp <- c()


f <- function(vol, spot=Spot, k){
  GBSPrice(PutCall='c', S=spot, K=k, T=T, r=r, b=b, sigma=vol)
}

Vega <- function(f, vol, spot=Spot, k) {
  d_vol <- 10e-5
  return( (f(vol+d_vol*vol, Spot, k)-f(vol-d_vol*vol, Spot, k))/(2*d_vol) )
}

Vanna <- function(f, vol, spot=Spot, k) {
  d_vol <- 10e-5
  d_spot <- 10e-5
  return( (f(vol+d_vol*vol, Spot+d_spot*Spot, k)+
           f(vol-d_vol*vol, Spot-d_spot*Spot, k)-
           f(vol+d_vol*vol, Spot-d_spot*Spot, k)-
           f(vol-d_vol*vol, Spot+d_spot*Spot, k))/(4*d_vol*d_spot) )
}
  
Volga <- function(f, vol, spot=Spot, k) {
  d_vol <- 10e-5
  return( (f(vol+d_vol*vol, Spot, k) - 2*f(vol, Spot, k) + f(vol-d_vol*vol, Spot, k))/d_vol**2 )
  }

vol.f <- function(f, vol = VolData.ATM, spot=Spot, k) {
  b = c(1:3)
  b[1] = Vega(f, vol, spot=Spot, k)
  b[2] = Vanna(f, vol, Spot, k)
  b[3] = Volga(f, vol, Spot, k)
  A = matrix(data = c(B.vega.benchmark, B.vanna.benchmark, B.volga.benchmark), nrow =3)
  A <- t(A)
  X = solve(A, b)
  return(VolData[[1]][2]*X[1] + VolData[[2]][2]*X[2] + VolData[[3]][2]*X[3])
}


for (i in 1:121) {
  vol.imp = append(vol.imp, vol.f(f, vol = VolData.ATM,spot = Spot, k = K[i]))
}

plot(K, vol.imp)

```


# Pricing a digital call

Recall that a digital call with strike $K$ pays one euro if $S_T \geq K$, and nothing otherwise.

Using the same logic as in the previous question, price a digital call, maturity $T=1$, struck at $K=105$. 

```{r, echo=F}
Vega <- function(f, vol, spot=Spot) {
  d_vol <- 10e-5
  return( (f(vol+d_vol*vol, Spot)-f(vol-d_vol*vol, Spot))/(2*d_vol) )
}

Vanna <- function(f, vol, spot=Spot) {
  d_vol <- 10e-5
  d_spot <- 10e-5
  return( (f(vol+d_vol*vol, Spot+d_spot*Spot)+
           f(vol-d_vol*vol, Spot-d_spot*Spot)-
           f(vol+d_vol*vol, Spot-d_spot*Spot)-
           f(vol-d_vol*vol, Spot+d_spot*Spot))/(4*d_vol*d_spot) )
}
  
Volga <- function(f, vol, spot=Spot) {
  d_vol <- 10e-5
  return( (f(vol+d_vol*vol, Spot) - 2*f(vol, Spot) + f(vol-d_vol*vol, Spot))/d_vol**2 )
  }
```

```{r, echo=TRUE}
r<-0
b<-0
T<-1
Spot <- 100
K <- 105

epsi <- 0.000001


VolData.ATM <- VolData[[2]][2]

f <- function(vol=VolData.ATM, spot=Spot){
  GBSPrice(PutCall='c', S=spot, K=K-epsi, T=T, r=r, b=b, sigma=vol)
}
B.vega <- Vega(f, VolData.ATM, Spot)
B.vanna <- Vanna(f, VolData.ATM, Spot)
B.volga <- Volga(f, VolData.ATM, Spot)
print("vega for K=105 - espilon")
B.vega
print("vanna for K=105 - espilon")
B.vanna
print("volga for K=105 - espilon")
B.volga
b.risk.moins <- c(B.vega, B.vanna, B.volga)
X.moins <- solve(A, b.risk.moins)

O.BS.moins <- GBSPrice(PutCall='c', S=100, K=K-epsi, T=T, r=r, b=b, sigma=VolData[[2]][2])


somme <- 0
for (i in 1:3) {
  somme <- somme + X.moins[i]*(C.M[i]-C.BS[i])
}
O.M.moins <- O.BS.moins + somme

print("price with BS for K=105 - espilon")
print(O.BS.moins)
print("price ajusted for K=105 - espilon")
print(O.M.moins)

```

```{r, echo=F}
VolData.ATM <- VolData[[2]][2]
##############
f <- function(vol=VolData.ATM, spot=Spot){
  GBSPrice(PutCall='c', S=spot, K=K+epsi, T=T, r=r, b=b, sigma=vol)
}
B.vega <- Vega(f, VolData.ATM, Spot)
B.vanna <- Vanna(f, VolData.ATM, Spot)
B.volga <- Volga(f, VolData.ATM, Spot)
print("vega for K=105 - espilon")
B.vega
print("vanna for K=105 - espilon")
B.vanna
print("volga for K=105 - espilon")
B.volga
b.risk.plus <- c(B.vega, B.vanna, B.volga)
X.plus <- solve(A, b.risk.plus)

O.BS.plus <- GBSPrice(PutCall='c', S=100, K=K+epsi, T=T, r=r, b=b, sigma=VolData[[2]][2])


somme <- 0
for (i in 1:3) {
  somme <- somme + X.plus[i]*(C.M[i]-C.BS[i])
}
O.M.plus<- O.BS.plus + somme

print("price with BS for K=105 + espilon")
print(O.BS.plus)
print("price ajusted for K=105 + espilon")
print(O.M.plus)

```


```{r, echo=F}
print("price for a digital call paying 2*epsi")
O =O.M.moins - O.M.plus
print(O)


prix.call.digital = O*(1/(2*epsi))
print("price for a digital call paying 1")
prix.call.digital
```

